---
title: "Pair Trading of GLD and GDX"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Example 3.6: 
#### Pair Trading of GLD and GDX

GLD versus GDX is a good candidate for pair trading because GLD reflects the spot price of gold, and GDX is a basket of gold-mining stocks. It makes intuitive sense that their prices should move in tandem.

(also see, e.g., epchan.blogspot.com/2006/11/reader-suggested-possible-trading.html)


We perform a regression analysis on the training set to determine the hedge ratio between GLD and GDX, and then define entry and exit thresholds for a pair-trading strategy.

We will see how optimiz- ing these thresholds on the training set changes the performance on the test set.


```{r}
library(quantmod)
smbls <- c("GLD", "GDX")
getSymbols(smbls, auto.assign = T)
```

```{r}
n <- length(GDX$GDX.Adjusted)
upper <- round(length(GDX$GDX.Adjusted)*.8)
index(GDX[upper,])
train <- 1:upper
test <- (upper+1):n

gold_prices <- merge(GLD$GLD.Adjusted, GDX$GDX.Adjusted)
colnames(gold_prices) <- smbls

gold_lret <- diff(log(gold_prices))
gold_lret_train <- gold_lret[train,]
gold_lret_test <- gold_lret[test,]

head(gold_lret_train)
head(gold_lret_test)
```

```{r}
plot.zoo(gold_prices, plot.type = "single", col=2:3, ylab = "price", xlab = "time", main = "Price over Time")
legend("topleft", smbls, pch = 10, col = 2:3)

plot.zoo(gold_lret, plot.type = "multiple", col=2:3, xlab = "time", main = "Log-Returns over Time")
```

```{r}
results <- lm(gold_prices$GLD ~ gold_prices$GDX)
hedge_ratio <- results$coefficients[1]
spread=gold_prices$GLD-hedge_ratio*gold_prices$GDX

plot.zoo(spread[train,])
plot.zoo(spread[test,])

spread_mean <- mean(spread[train,]);
spread_sd <- sd(spread[train,]);
zscore <- (spread - spread_mean)/spread_sd;

# buy spread when its value drops below 2 standard deviations.
longs <- zscore <= -1

# short spread when its value rises above 2 standard deviations.
shorts <- zscore >= 1
# exit any spread position when its value is within 1 standard deviation of its mean. 
exits <- abs(zscore) <= 0.5

# initialize positions array 
pos <- xts(rep(NA, n), order.by = index(gold_prices))
positions <- merge(pos, pos)
colnames(positions) <- smbls

# long entries
positions[shorts, ] <- matrix(rep(c(-1,1), length(which(shorts))), ncol =2, byrow = T)
# short entries
positions[longs,  ] <- matrix(rep(c(-1,1), length(which(longs))), ncol =2, byrow = T)
# View(positions["2010-11/12"])
# View(longs["2010-11/12"])

# exit positions
positions[exits,  ] <- matrix(rep(c(0, 0), length(which(exits))), ncol =2, byrow = T)

positions = na.locf(positions)
```


```{r}
daily_ret <- (gold_prices - lag(gold_prices))/lag(gold_prices)
daily_ret <- daily_ret[-1]

plot.zoo(daily_ret, plot.type = "multiple", col=2:3, xlab = "time", main = "Log-Returns over Time")

pnl <- lag(positions)*daily_ret

# the Sharpe ratio on the training set should be about 2.3
sharpe_train <- sqrt(252)*mean(pnl[train[-1],]/sd(pnl[train[-1],]))
sharpe_train


# the Sharpe ratio on the test set should be about 1.5
sharpe_testset=sqrt(252)*mean(pnl[test-1,])/sd(pnl[test-1,])
plot.zoo(cumsum(pnl[test-1,]), plot.type = "single", col=2:3, ylab = "price", xlab = "time", main = "Price over Time")
legend("topleft", smbls, pch = 10, col = 2:3)
sharpe_testset
```

